Cluster Dynamics with HETDEX at z < 0.5 - I: Simulated Performance, Mass Recovery, and Limits to Cosmology

### 2016-05-18 paper1.tex
more casey comments

\begin{abstract}
\noindent
The distribution of massive clusters of galaxies depends strongly on the total cosmic mass density, the mass variance, and the dark energy equation of state. As such, measures of galaxy clusters can provide constraints on these parameters and even test models of gravity, but only if observations of clusters can lead to accurate estimates of their total masses. Here, we carry out a study to investigate the ability of a blind spectroscopic survey to recover accurate galaxy cluster masses through their velocity dispersion using probability based and machine learning methods. We focus on the Hobby Eberly Telescope Dark Energy Experiment (HETDEX), which will employ new Visible Integral-Field Replicable Unit Spectrographs (VIRUS), over 420 \degsq\ on the sky with a 1/4.5 fill factor. VIRUS covers the blue/optical portion of the spectrum ($3500-5500~\AAA$), allowing surveys to measure redshifts for a large sample of galaxies out to $z < 0.5$ based on their absorption features or [\ion{O}{ii}] $\lambda$3727 emission (and Lyman-$\alpha$ over $1.9 < z < 3.5$). We use a detailed mock galaxy catalog from a semi-analytic model to simulate surveys observed with VIRUS, including: (1) a blind, HETDEX-like survey with an incomplete but uniform spectroscopic selection function; and (2) IFU surveys that target clusters directly, obtaining spectra of all galaxies in a VIRUS-sized field. For both surveys, we include realistic uncertainties from galaxy magnitude and line-flux limits. We benchmark both surveys against spectroscopic observations with ``perfect" knowledge of galaxy line-of-sight velocities. We find that with HETDEX we can recover cluster masses to $\sim0.1$ dex which can be further improved to $<0.1$ dex with targeted followup observations. This level of cluster mass recovery enables us to place $<20$\% constraints on $\sigma_8$, and leverage the unique properties of the observations to provide important calibrations for the optical richness-cluster mass relation.
\end{abstract}
% Using both observing strategies, we produce cluster mass probability density functions $P(M_{200c}|\vec{x})$ of a cluster mass ($M_{200c}$) as a function of an observable parameter or set of observable parameters, $\vec{x}$, where we consider cases of where $\vec{x} = \{\sigma\}$, $=\{\sigma, z\}$, and $=\{\sigma,z, N_{gal}\}$ for velocity dispersion, $\sigma$, redshift, and number of observed galaxies, $N_{gal}$.
% \end{abstract}

@@

As mass is not a direct observable, a lot of work is underway to characterize galaxy cluster masses with an observable feature of galaxy clusters. The goal is to constrain, as best possible, $P(M|z, \vec{x})$ the probability ($P$) that a galaxy cluster of given mass ($M$), located at redshift ($z$) determined using an observable parameter or parameters ($\vec{x}$). Generally, cluster mass calibrations fall into two distinct camps, simulation based and direct or statistical calibration.

One could use various simulations to attempt to calibrate this observable--mass relation \citeeg{Vanderlinde2010, Sehgal2011}. However, the primary challenge to this method is the incomplete understanding of the baryonic physics which take place in galaxy cluster environments. While there have been (and continue to be) many improvements in the accuracy and power of simulations it is doubtful that in the coming years they will reach the accuracy level required where the observable--mass relation is dominated only by statistics \citep{Weinberg2013}.

The second broad camp is the direct calibration of cluster masses. This recipe has two distinct but not always independent tracks. The ``direct'' method uses direct observations of a small set of clusters and then uses known mass estimators, including X-ray temperatures and luminosities \citeeg{Mantz2010, Rykoff2014}, microwave observations \citeeg{Vanderlinde2010, Sehgal2011}, optical richness \citeeg{Abell1958, Rykoff2012} or weak lensing (WL; \eg\ \citealt{Rozo2010}) as examples, which provide a ``true'' mass. This directly calibrates the observable--mass relation which is then applied to a much larger sample. The complications lie in that the ``true'' masses are, in fact, estimations, and the methods used to recover these cluster masses are subject to their own limitations. X-ray based cluster masses assume hydrostatic equilibrium \citeeg{Mantz2015} which may only be valid for a very small number and range of cluster masses. The Sunyaev-Zel'dovich Effect (SZE; \citealt{Sunyaev1972}), which uses the up-scattering of cosmic microwave background (CMB) photons to estimate cluster masses, provides accurate estimations of mass, but the ability to detect low mass galaxy clusters is currently limited by technology \citeeg{Carlstrom2002a}. WL estimates are, in principle, correct in the mean, but they suffer from signal-to-noise requirements, limiting their usefulness in low mass clusters (where the lensing signal is particularly weak), and potentially suffer from line-of-sight effects as WL is sensitive to all mass along the line-of-sight. Virial mass estimators which determine the cluster mass based on the motions of the member galaxies \citeeg{Ruel2014, Sifon2015} are promising in that it is a direct measurement of the depth of clusters potential well, but suffer from systematics due to cluster formation physics which disrupts the velocity field.

@@

Large-area sky surveys, both on going and planned, are revolutionizing cluster cosmology using a large range of wavelengths. The South Pole Telescope (SPT; \citealt{Carlstrom2011}) and the Atacama Cosmology Telescope (ACT; \citealt{Swetz2011}) are discovering many clusters through the SZE. Optically, the on going Dark Energy Survey (DES; \citealt{DES2005}) and planned Large Synoptic Survey Telescope (LSST; \citealt{LSST2012}) will identify many thousands of clusters to much lower masses than is possible with SZE measurements. However, regardless of the discovery method used, spectroscopic followup is needed to further constrain $P(M|z,X)$. But as the cluster dataset grows to many tens of thousands of clusters individual followup becomes increasingly impractical. Therefore, large spectroscopic surveys are needed to more fully constrain the observable--mass relation of clusters.

@@

We begin in Section \ref{sec:Data} by giving an overview of what data is used, how it is created, and how we make our ``observations.'' Details about the determination of cluster parameters, velocity dispersion, total mass, etc., are discussed in Section \ref{sec:recovery}. Next, we present the results of our study in Section \ref{sec:results} and discuss their implications in Section \ref{sec:discussion}. Finally we summarize our findings in Section \ref{sec:summary}. A followup to this work \citep{Boada2016a} will investigate how the techniques developed here will work in practice.

@@

\subsection{The Buzzard Mock Catalogs}
\begin{figure*} 
	\includegraphics[width=\textwidth]{figures/oii_sdss.pdf} 
	\caption{\textit{Left}: CMD of 503,113 $z<0.2$ galaxies take from the SDSS DR12 where the shading scales with the density of points. The two colored boxes show regions containing potential catalog galaxies. \textit{Right}: Probability histograms of the Log [\ion{O}{ii}] luminosity for the SDSS galaxies located in the two highlighted regions on the right. The [\ion{O}{ii}] luminosities are assigned to catalog galaxies from slice sampling the probability histogram and converted to fluxes using the redshift of each galaxy.} \label{fig: oii sdss} 
\end{figure*}

The Buzzard mock galaxy catalogs cover 398.49 \degsq\ between $4^h< RA < 6^h$ and $-61\degree < DEC < -41\degree$ and are derived from a combination of Sub-halo Abundance Matching (ShAM) and ADDSEDs (Adding Density Dependent Spectral Energy Distributions) tied to an in house n-body cosmological simulation. A brief description of the catalog creation is described as follows. The initial conditions are generated with a second-order Lagrangian perturbation theory using {\sc 2LPTic} \citep{Crocce2006}. Dark matter (DM) n-body simulations are run using {\sc LGadget-2} (a version of {\sc Gadget-2}; \citealt{Springel2005}). The DM halos are identified using the {\sc ROCKSTAR} halo finder \citep{Behroozi2013} which also calculates halo masses.

@@

\subsection{Conditional [\ion{O}{ii}] Flux Probability Distribution Functions}\label{sec: oii luminosity}
We use the SDSS DR12 \citep{Alam2015} catalogs to assign [\ion{O}{ii}] emission line strengths to the galaxies in the Buzzard catalog. We use 503,113 galaxies from $z = 0.05 - 0.2$, which are selected with no redshift warning, and place each galaxy on a color-magnitude diagram (CMD) of $M_r$ and $g-r$; see the left panel of Figure \ref{fig: oii sdss}.

@@

Figure~\ref{fig: oii sdss} illustrates this process. The numbered boxes in the left panel show two bins ($M_r, g-r = -17.7,~0.49$ and $M_r, g-r = -21.4,~1.24$) in the CMD of $0.05 < z < 0.2$ SDSS galaxies. The right panel shows the Log [\ion{O}{ii}] luminosity distribution functions, $p($[\ion{O}{ii}]$| M_r, g-r)$, which we use to assign [\ion{O}{ii}] luminosity to each object. The luminosity is then converted into an [\ion{O}{ii}] flux through 
\begin{equation}
	F = \frac{L}{4\pi D_L}
\end{equation}
where $D_L$ is the luminosity distance \citeeg{Hogg1999}.

\subsection{HETDEX}\label{sec: hetdex}
\begin{figure} 
	\includegraphics[width=\columnwidth]{figures/pointings.pdf} 
	\caption{Representative observation tiling scheme for the HETDEX $16' \times 16'$ pointings. Each colored square is a single VIRUS IFU and the dashed octagons approximate the size of a single observation. See the text for more details.} \label{fig: ifu layout} 
\end{figure}

\editorial{There are quite a few words copied from Casey's SHELA paper.}
HETDEX will perform a blind survey which will measure the redshifts of $8 \times 10^5$ LAE galaxies between $1.9 < z < 3.5$ using a collection of 78 wide-field IFUs spectrographs covering the wavelength region $3500 - 5500$~\AAA\ at R $\sim$ 750 \citep{Hill2008}. The goal of these observations will be to provide sub-percent level measurements of the Hubble expansion parameter and the angular diameter distance at $z \sim 2$ via the large scale distribution of galaxies in the redshift range of HETDEX. The result will be a significant constraint on the evolution of dark energy that is competitive with (and independent of) values based on surveys of the Ly$\alpha$ forest.

@@

\subsection{Mock Observations}\label{sec: observations}
This means that when selecting galaxies from the Buzzard catalog we assume an observation for all galaxies laying within a turquoise, IFU square in Figure~\ref{fig: ifu layout}. Galaxies which lie between the IFUs are missed, as well as the galaxies which lie between the pointings, as there is no overlap between one pointing and the next. To cover the 398.49 \degsq\ field of the Buzzard catalog we require 5370 pointings where 0.015 \degsq\ of each pointing is covered by an IFU. The total area of the sky covered by an IFU is 80.80 \degsq\ which gives a filling factor of 1/4.65, slightly decreased from the expected filling factor of 1/4.5.

In this work we consider two separate observing strategies, Targeted and Survey. The Targeted observations use ``direct'' observations where each cluster is targeted individually, and every cluster member galaxy is assumed to be observed. The Survey observations mimic the HETDEX observation pattern across the sky, where no cluster is directly targeted and not all cluster member galaxies are observed. Both of these observations have HETDEX-like galaxy detection thresholds, and for comparison we also include a set of Targeted observations with ``perfect'' knowledge. Perfect observations assume no detection threshold, if a cluster member galaxy is observed, it is also detected. This provides an important best-case scenario, and differs from the true cluster properties because the cluster properties are still calculated. These observations provide three levels of quality with ``Perfect knowledge'' being the highest and Survey being the lowest.

@@

In simple terms, the cluster redshift is the mean of the redshifts of all galaxies associated with the cluster, where the mean is the first moment of the velocity (redshift) distribution function $p(z)$. In practice, the first moment is strongly subject to outliers, so we rely instead on the biweight location estimator \citep{Beers1990} through\footnote{Implemented as part of the {\sc astLib} Python library. See \hbox{\url{http://astlib.sourceforge.net}}}:
\begin{equation}
C_{BI} = M + \frac{\sum_{|u_i| < 1} (x_i - M)(1-u_i^2)^2}{\sum_{|u_i| < 1}(1-u_i^2)^2}
\end{equation}
where $x_i$ are the individual redshifts, $M$ is the median redshift and $u_i$ is given by:
\begin{equation}
	u_i = \frac{(x_i - M)}{C\,\mathrm{MAD}}.
\end{equation}
MAD is the median absolute deviation, also defined in \cite{Beers1990}, and $C$ is the a tuning constant. We choose $C=6$ (the suggested value) which balances computational speed and location accuracy.

@@

\subsection{Line-of-Sight Velocity Dispersion}\label{sec: LOSVD}
We calculate the LOSV to each galaxy, where
\begin{equation}
	\mathrm{LOSV} = c\frac{z - z_c}{1+z_c}
\end{equation}
and $c$ is the speed of light in \kms, $z$ is the redshift of the individual galaxy, and $z_c$ is the overall cluster redshift described in the previous subsection.

We follow the maximum likelihood method of \cite{Walker2006} to estimate the line-of-sight velocity dispersion (LOSVD). We maximize the probability function 
\begin{equation}\label{eq: jointGaussian}
p(\{v_1, ..., v_N\})=\displaystyle\prod_{i=1}^{N}\frac{1}{\sqrt{2\pi(\sigma_i^2+\sigma_{1D}^2)}}\exp\biggl[-\frac{1}{2}\frac{(v_i-\langle u \rangle)^2}{(\sigma_i^2+\sigma_{1D}^2)}\biggr]
\end{equation}
where $\sigma_{1D}$, $\langle\mu\rangle$, and $\sigma_i$ is the LOSVD, the average radial velocity and the error on the individual LOSVs (which we have assumed to be normally distributed) respectively, using a Monte Carlo Markov Chain (MCMC) sampler ({\sc emcee}\footnote{\url{http://dan.iel.fm/emcee/current/}}; \citealt{Foreman-Mackey2013}) which is based on affine-invariant ensemble sampler (see \citealt{Goodman2010} for details on affine-invariant samplers). We draw twenty thousand samples from the posterior probability distribution using simple priors, $\langle\mu\rangle$ lies between the maximum and minimum LOSV and $0< \sigma_{1D} < 1400$ \kms. We set the upper limit on the LOSVD as the LOSVD corresponding to a $10^{16}$ \Msol\ cluster at $z\sim0.0$, higher mass than any expected cluster in Buzzard. When the full distribution of LOSVDs is not used, the final LOSVD is quoted as the median value of the posterior probability distribution with 68\% error bars defined as the second momment of the same distribution.

@@

\subsubsection{Other Estimates of Dynamical Mass -- Introduction}
In the following sections we use two methods to predict the mass of a cluster based on other observables. Often the cluster mass is estimated based on a single observable, X-ray temperature, LOSVD, richness and others (see Section~\ref{sec: Introduction} for referenced examples). Here we combine many observables to attempt to correct the mass inferred solely from the velocity dispersion. The first method is traditional ``probability based'' where we marginalize over a series of observables to find the most probable cluster mass. The second is based on a machine learning (ML) algorithm which attempts to infer the relationship between the observables and the desired output, the cluster mass. Both of these methods are examples of supervised learning algorithms where the relationship between the observable parameters and the target parameter (the cluster mass) are both known.

@@

In this method, the cluster masses are predicted much in the same way we assign [\ion{O}{ii}] flux to the Buzzard catalog galaxies. Figure~\ref{fig: probability corner} illustrates the process. The left panel shows the two dimensional (joint probability) projections of the posterior probability distributions of the feature training data. The conditional probability of the cluster mass $P(M_{200c}|\vec{x}= \{ x_1,x_2,...\})$ (shown in the right panel) is determined by selecting a region in the joint probability distribution. For example, using the LOSVD and redshift features we create $P(M_{200c}|\vec{x})$ for two test galaxies, shown by the green and pink boxes in the left panel of Figure~\ref{fig: probability corner}. These example galaxies have features $\vec{x} = \{\sigma=200\, \kms,z=0.16\}$ and $\vec{x} = \{\sigma=400\, \kms,z=0.36\}$. We select all galaxies in our training sample with similar features and create the conditional probability distributions shown in the right panel.

For the clusters making up the \emph{test} sample the mass is unknown (it is what we are trying to predict) but the other features are known. To determine the mass probability distribution of a test cluster, $P(M_{200c})$, we combine the conditional probability distribution, $P(M_{200c}|\vec{x})$, created previously with the probability distribution of $\sigma$, the LOSVD, through Equation~\ref{eq: Pm}.
\begin{equation}\label{eq: Pm}
	P(M_{200c}) = \int P(M_{200c}|\vec{x})\ P(\sigma)\ d\sigma
\end{equation}
The expected mass is determined by calculating the first moment of the probability density. This becomes our ``predicted'' cluster mass, $M_{pred}$.
\begin{equation}\label{eq: expected mass}
	M_{pred}= \int M_{200c}\ P(M_{200c})\ dM_{200c}
\end{equation}
The confidence interval associated with this prediction can be estimated two ways. First, by calculating the second moment of the probability density through
\begin{equation}\label{eq: variance}
	\mathrm{Var} = \int (M_{200c} - M_{pred})^2\ P(M_{200c})\ dM_{200c}
\end{equation}
or by drawing many samples from $P(M_{200c})$ and calculating the values at the 16th and 84th percentile. In practice we find that both methods produce similar results for a large number of trials. Therefore, we quote predicted masses as the most probable mass given by Equation~\ref{eq: expected mass} and associated 68\% error estimated through Equation~\ref{eq: variance}.

@@

Here we use an averaging ensemble learning method known as a forest of randomized decision trees, often shorten to just random forest (RF; \citealt{Ho1995, Ho1998}). Decision trees can be visualized a flow chart where forks are the branches of the tree. The path along the tree is decided by the values of the feature(s) at each branch. RF estimators use a random subset of the training set at each fork to decide which path should be followed. The final prediction is then the average of all the final predictions from the trees. We use RF regression methods as implemented in {\sc Scikit-Learn} \citep{Pedregosa2012}.

>>>>

\begin{abstract}
\noindent
The distribution of massive clusters of galaxies depends strongly on the total cosmic mass density, the mass variance, and the dark energy equation of state. As such, measures of galaxy clusters can provide constraints on these parameters and even test models of gravity, but only if observations of clusters can lead to accurate estimates of their total masses. Here, we carry out a study to investigate the ability of a blind spectroscopic survey to recover accurate galaxy cluster masses through their velocity dispersion using probability based and machine learning methods. We focus on the Hobby Eberly Telescope Dark Energy Experiment (HETDEX), which will employ new Visible Integral-Field Replicable Unit Spectrographs (VIRUS), over 420 \degsq\ on the sky with a 1/4.5 fill factor. VIRUS covers the blue/optical portion of the spectrum ($3500-5500~\AAA$), allowing surveys to measure redshifts for a large sample of galaxies out to $z < 0.5$ based on their absorption features or [\ion{O}{ii}] $\lambda$3727 emission (and Lyman-$\alpha$ over $1.9 < z < 3.5$). We use a detailed mock galaxy catalog from a semi-analytic model to simulate surveys observed with VIRUS, including: (1) a blind, HETDEX-like survey with an incomplete but uniform spectroscopic selection function; and (2) IFU surveys that target clusters directly, obtaining spectra of all galaxies in a VIRUS-sized field. For both surveys, we include realistic uncertainties from galaxy magnitude and line-flux limits. We benchmark both surveys against spectroscopic observations with ``perfect" knowledge of galaxy line-of-sight velocities. With HETDEX, we can recover cluster masses to $\sim0.1$ dex which can be further improved to $<0.1$ dex with targeted follow-up observations. This level of cluster mass recovery enables constraints on $\sigma_8$ to $<20$\%, and the unique properties of the observations will provide important calibrations for the optical richness-cluster mass relation.
\end{abstract}
% Using both observing strategies, we produce cluster mass probability density functions $P(M_{200c}|\vec{x})$ of a cluster mass ($M_{200c}$) as a function of an observable parameter or set of observable parameters, $\vec{x}$, where we consider cases of where $\vec{x} = \{\sigma\}$, $=\{\sigma, z\}$, and $=\{\sigma,z, N_{gal}\}$ for velocity dispersion, $\sigma$, redshift, and number of observed galaxies, $N_{gal}$.
% \end{abstract}

@@

As mass is not a direct observable, a lot of work is underway to characterize galaxy cluster masses with an observable feature of galaxy clusters. The goal is to constrain $P(M|z, \vec{x})$ the probability density ($P$) that a galaxy cluster of given mass ($M$), located at redshift ($z$) determined using an observable parameter or parameters ($\vec{x}$). Generally, cluster mass calibrations are done in one of two ways, through simulations or direct or statistical calibration.

One could use various simulations to attempt to calibrate this observable-mass relation \citeeg{Vanderlinde2010, Sehgal2011}. However, the primary challenge to this method is the incomplete understanding of the baryonic physics which take place in galaxy cluster environments. While there have been (and continue to be) many improvements in the accuracy and power of simulations it is doubtful that in the coming years they will reach the accuracy level required where the observable--mass relation is dominated only by statistics \citep{Weinberg2013}.

The second broad method is the direct calibration of cluster masses. This recipe has two distinct but not always independent tracks. The ``direct'' method uses observations of a relatively small set of clusters and then uses known mass estimators, including X-ray temperatures and luminosities \citeeg{Mantz2010, Rykoff2014}, microwave observations \citeeg{Vanderlinde2010, Sehgal2011}, optical richness \citeeg{Abell1958, Rykoff2012} or weak lensing (WL; \eg\ \citealt{Rozo2010}) as examples, which provide a ``true'' mass. This directly calibrates the observable-mass relation which is then applied to a much larger sample. The complications lie in that the ``true'' masses are, in fact, estimations, and the methods used to recover these cluster masses are subject to their own limitations. X-ray based cluster masses assume hydrostatic equilibrium \citeeg{Mantz2015} which may only be valid for a very small number and range of cluster masses. The Sunyaev-Zel'dovich Effect (SZE; \citealt{Sunyaev1972}), which uses the up-scattering of cosmic microwave background (CMB) photons to estimate cluster masses, provides accurate estimations of mass, but the ability to detect low mass galaxy clusters is currently limited by technology \citeeg{Carlstrom2002a}. WL estimates are, in principle, correct in the mean, but they suffer from signal-to-noise requirements, limiting their usefulness in low mass clusters (where the lensing signal is particularly weak), and potentially suffer from line-of-sight effects as WL is sensitive to all mass along the line-of-sight. Virial mass estimators which determine the cluster mass based on the motions of the member galaxies \citeeg{Ruel2014, Sifon2015} are promising in that it is a direct measurement of the depth of clusters potential well, but suffer from systematics due to cluster formation physics which disrupts the velocity field.

@@

Large-area sky surveys, both on going and planned, are revolutionizing cluster cosmology using a large range of wavelengths. The South Pole Telescope (SPT; \citealt{Carlstrom2011}) and the Atacama Cosmology Telescope (ACT; \citealt{Swetz2011}) are discovering many clusters through the SZE. Optically, the on going Dark Energy Survey (DES; \citealt{DES2005}) and planned Large Synoptic Survey Telescope (LSST; \citealt{LSST2012}) will identify many thousands of clusters to much lower masses than is possible with SZE measurements. However, regardless of the discovery method used, spectroscopic follow-up is needed to further constrain $P(M|z,X)$. But as the cluster dataset grows to many tens of thousands of clusters individual follow-up becomes increasingly impractical. Therefore, large spectroscopic surveys are needed to more fully constrain the observable--mass relation of clusters.

@@

We begin in Section \ref{sec:Data} by giving an overview of what data is used, how it is created, and how we make our ``observations.'' Details about the determination of cluster parameters, velocity dispersion, total mass, etc., are discussed in Section \ref{sec:recovery}. Next, we present the results of our study in Section \ref{sec:results} and discuss their implications in Section \ref{sec:discussion}. Finally we summarize our findings in Section \ref{sec:summary}. A follow-up to this work \citep{Boada2016a} will investigate how the techniques developed here will work in practice.

@@

\subsection{The Buzzard Mock Catalogs}
The Buzzard mock galaxy catalogs cover 398.49 \degsq\ between $4^h< RA < 6^h$ and $-61\degree < DEC < -41\degree$ and are derived from a combination of Sub-halo Abundance Matching (ShAM) and ADDSEDs (Adding Density Dependent Spectral Energy Distributions) tied to an in house n-body cosmological simulation. A brief description of the catalog creation is described as follows. The initial conditions are generated with a second-order Lagrangian perturbation theory using {\sc 2LPTic} \citep{Crocce2006}. Dark matter (DM) n-body simulations are run using {\sc LGadget-2} (a version of {\sc Gadget-2}; \citealt{Springel2005}). The Buzzard catalogs adopt the following cosmological parameters: $\Omega_m = 0.286$, $\Omega_\Lambda = 0.714$, $H_0 = 70$ \kms \mpc, $\sigma_8 = 0.82$, and $n_s = 0.96$. The DM halos are identified using the {\sc ROCKSTAR} halo finder \citep{Behroozi2013} which also calculates halo masses.

@@

\subsection{Conditional [\ion{O}{ii}] Flux Probability Distribution Functions}\label{sec: oii luminosity}
\begin{figure*} 
	\includegraphics[width=\textwidth]{figures/oii_sdss.pdf} 
	\caption{\textit{Left}: CMD of 503,113 $z<0.2$ galaxies take from the SDSS DR12 where the shading scales with the density of points. The two colored boxes show regions containing potential catalog galaxies. \textit{Right}: Probability histograms of the Log [\ion{O}{ii}] luminosity for the SDSS galaxies located in the two highlighted regions on the right. The [\ion{O}{ii}] luminosities are assigned to catalog galaxies from slice sampling the probability histogram and converted to fluxes using the redshift of each galaxy.} \label{fig: oii sdss} 
\end{figure*}
We use the SDSS DR12 \citep{Alam2015} catalogs to assign [\ion{O}{ii}] emission line strengths to the galaxies in the Buzzard catalog. We use 503,113 objects classified as galaxies selected over $z = 0.02 - 0.2$ with {\sc zwarning $=0$} and a measured [\ion{O}{ii}] line flux signal-to-noise of five. Figure~\ref{fig: oii sdss} shows the color-magnitude diagram (CMD) of $M_r$ and $g-r$ for these galaxies.

@@

Figure~\ref{fig: oii sdss} illustrates this process. The numbered boxes in the left panel show the bins corresponding to two example Buzzard galaxies ($M_r, g-r = -17.7,~0.49$ and $M_r, g-r = -21.4,~1.24$). The right panel shows the Log [\ion{O}{ii}] luminosity distribution functions, $p($[\ion{O}{ii}]$| M_r, g-r)$, which we use to assign [\ion{O}{ii}] luminosity to each object. The luminosity is then converted into an [\ion{O}{ii}] flux through 
\begin{equation}
	F = \frac{L}{4\pi D_L}
\end{equation}
where $D_L$ is the luminosity distance \citeeg{Hogg1999}.

\subsection{HETDEX}\label{sec: hetdex}
\begin{figure} 
	\includegraphics[width=\columnwidth]{figures/pointings.pdf} 
	\caption{Representative observation tiling scheme for the HETDEX $16' \times 16'$ pointings. Each turqoise square represents the position of a single VIRUS IFU, and the dashed octagons approximate the size of a single observation. Inside each IFU HETDEX will achieve near complete coverage through three dithers. See the text for more details.} \label{fig: ifu layout} 
\end{figure}

We designed the results study here to be used with HETDEX, a large blind spectroscopic survey. HETDEX will measure the redshifts of $8 \times 10^5$ LAE galaxies between $1.9 < z < 3.5$ using a collection of 78 wide-field IFU spectrographs covering the wavelength region $3500 - 5500$~\AAA\ at R $\sim$ 750 \citep{Hill2008}. The goal of these observations will be to provide sub-percent level measurements of the Hubble expansion parameter and the angular diameter distance at $z \sim 2$ via the large scale distribution of galaxies in the redshift range of HETDEX. The result will be a significant constraint on the evolution of dark energy that is competitive with (and independent of) values based on surveys of the Ly$\alpha$ forest.

@@

\subsection{Mock Observations}\label{sec: observations}
When selecting galaxies from the Buzzard catalog we assume an observation for all galaxies laying within a turquoise, IFU square in Figure~\ref{fig: ifu layout}. In practice, this is achieved by three dither positions at each pointing. Galaxies which lie between the IFUs are missed, as well as the galaxies which lie between the pointings, as there is no overlap between one pointing and the next. To cover the 398.49 \degsq\ field of the Buzzard catalog we require 5370 pointings where 0.015 \degsq\ of each pointing is covered by an IFU. The total area of the sky covered by an IFU is 80.80 \degsq\ which gives a filling factor of 1/4.65, slightly decreased from the expected filling factor of 1/4.5.

In this work we consider two separate observing strategies, Targeted and Survey. The Targeted observations use ``direct'' observations where each cluster is targeted individually, and every cluster member galaxy is assumed to be observed. The Survey observations mimic the HETDEX observation pattern across the sky, where no cluster is directly targeted and not all cluster member galaxies are observed. Both of these observations have HETDEX-like galaxy detection thresholds (described in the previous subsection), so while a galaxy may be observed, a redshift will only be measured if the galaxy satisfies the continuum brightness or emission line flux limits for the HETDEX survey. For comparison we also include a set of Targeted observations with ``perfect'' knowledge which assume no detection threshold, if a cluster member galaxy is observed, it is also detected. This provides an important best-case scenario, and differs from the true cluster properties between we are still calculating the cluster mass from the observed member galaxy velocity dispersions, and represent the best possible mass estimates using velocity dispersion as the observable. These observations provide three levels of quality with ``Perfect knowledge'' being the highest and Survey being the lowest.

@@

In simple terms, the cluster redshift is the mean of the redshifts of all galaxies associated with the cluster, where the mean is the first moment of the velocity (redshift) distribution function $p(z)$. In practice, the first moment is strongly subject to outliers, so we rely instead on the biweight location estimator \citep{Beers1990} through\footnote{Implemented as part of the {\sc astLib} Python library. See \hbox{\url{http://astlib.sourceforge.net}}}:
\begin{equation}
C_{BI} = M + \frac{\sum_{|u_i| < 1} (z_i - M)(1-u_i^2)^2}{\sum_{|u_i| < 1}(1-u_i^2)^2}
\end{equation}
where $z_i$ are the individual redshifts, $M$ is the median redshift and $u_i$ is given by:
\begin{equation}
	u_i = \frac{(z_i - M)}{C\,\mathrm{MAD}}.
\end{equation}
MAD is the median absolute deviation, also defined in \cite{Beers1990}, and $C$ is the a tuning constant. We choose $C=6$ (the suggested value) which balances computational speed and location accuracy.

@@

\subsection{Line-of-Sight Velocity Dispersion}\label{sec: LOSVD}
We calculate the LOSV to each galaxy, where
\begin{equation}
	\mathrm{LOSV} = c\frac{z - z_c}{1+z_c}
\end{equation}
and $c$ is the speed of light, $z$ is the redshift of the individual galaxy, and $z_c$ is the overall cluster redshift described in the previous subsection.

We follow the maximum likelihood method of \cite{Walker2006} to estimate the line-of-sight velocity dispersion (LOSVD). We maximize the probability function 
\begin{equation}\label{eq: jointGaussian}
p(\{v_1, ..., v_N\})=\displaystyle\prod_{i=1}^{N}\frac{1}{\sqrt{2\pi(\sigma_i^2+\sigma_{1D}^2)}}\exp\biggl[-\frac{1}{2}\frac{(v_i-\langle u \rangle)^2}{(\sigma_i^2+\sigma_{1D}^2)}\biggr]
\end{equation}
where $\sigma_{1D}$, $\langle\mu\rangle$, and $\sigma_i$ is the LOSVD, the average radial velocity and the error on the individual LOSVs (which we have assumed to be normally distributed) respectively, using a Monte Carlo Markov Chain (MCMC) sampler ({\sc emcee}\footnote{\url{http://dan.iel.fm/emcee/current/}}; \citealt{Foreman-Mackey2013}) which is based on affine-invariant ensemble sampler (see \citealt{Goodman2010} for details on affine-invariant samplers). We draw twenty thousand samples from the posterior probability distribution using simple priors, $\langle\mu\rangle$ lies between the maximum and minimum LOSV and $0< \sigma_{1D} < 1400$ \kms. We set the upper limit on the LOSVD as the LOSVD corresponding to a $10^{16}$ \Msol\ cluster at $z\sim0.0$, higher mass than any expected cluster in Buzzard. When the full distribution of LOSVDs is not used, the final LOSVD is quoted as the median value of the posterior probability distribution with 68\% error bars defined as the square root of second moment of the same distribution, the standard deviation.

@@

\subsubsection{Other Estimates of Dynamical Mass -- Introduction}
In the following subsections we use two methods to predict the mass of a cluster based on other observables. Often the cluster mass is estimated based on a single observable, X-ray temperature, LOSVD, richness and others (see Section~\ref{sec: Introduction} for referenced examples). Here we combine many observables to attempt to correct the mass inferred solely from the velocity dispersion. The first method is traditional ``probability based'' where we marginalize over a series of observables to find the most probable cluster mass. The second is based on a machine learning (ML) algorithm which attempts to infer the relationship between the observables and the desired output, the cluster mass. Both of these methods are examples of supervised learning algorithms where the relationship between the observable parameters and the target parameter (the cluster mass) are both known.

@@

In this method, the cluster masses are predicted using the method illustrated in Figure~\ref{fig: probability corner}. The left panel shows the two dimensional (joint probability) projections of the posterior probability distributions of the feature training data. The conditional probability of the cluster mass $P(M_{200c}|\vec{x}= \{ x_1,x_2,...\})$ (shown in the right panel) is determined by selecting a region in the joint probability distribution. For example, using the LOSVD and redshift features we create $P(M_{200c}|\vec{x})$ for two test galaxies, shown by the green and pink boxes in the left panel of Figure~\ref{fig: probability corner}. These example galaxies have features $\vec{x} = \{\sigma=200\, \kms,z=0.16\}$ and $\vec{x} = \{\sigma=400\, \kms,z=0.36\}$. We select all galaxies in our training sample with similar features and create the conditional probability distributions shown in the right panel.

For the clusters making up the \emph{test} sample the mass is unknown (it is what we are trying to predict) but the other features are known. To determine the mass probability distribution of a test cluster, $P(M_{200c})$, we combine the conditional probability distribution, $P(M_{200c}|\vec{x})$, created previously with the probability distribution of $\sigma$, the LOSVD, through Equation~\ref{eq: Pm}.
\begin{equation}\label{eq: Pm}
	P(M_{200c}) = \int P(M_{200c}|\vec{x})\ P(\sigma)\ d\sigma
\end{equation}
The expected mass is determined by calculating the first moment of the probability density. This becomes our ``predicted'' cluster mass, $M_{pred}$.
\begin{equation}\label{eq: expected mass}
	M_{pred}= \int M_{200c}\ P(M_{200c})\ dM_{200c}
\end{equation}
The confidence interval associated with this prediction can be estimated two ways. First, by calculating the second moment of the probability density through
\begin{equation}\label{eq: variance}
	\mathrm{Var} = \int (M_{200c} - M_{pred})^2\ P(M_{200c})\ dM_{200c}
\end{equation}
or by drawing many samples from $P(M_{200c})$ and calculating the values at the 16th and 84th percentile. In practice we find that both methods produce similar results for a large number of trials. Therefore, we quote predicted masses as the most probable mass given by Equation~\ref{eq: expected mass} and associated 68\% error estimated through the square root of Equation~\ref{eq: variance}.

@@

Here we use an averaging ensemble learning method known as a forest of randomized decision trees, often shorten to just random forest (RF; \citealt{Ho1995, Ho1998}). Decision trees can be visualized a flow chart where forks are the branches of the tree. The path along the tree is decided by the values of the feature(s) at each branch. RF estimators use a random subset of the training set at each fork to decide which path should be followed. The final prediction is then the mean of all the final predictions from the trees. We use RF regression methods as implemented in {\sc Scikit-Learn} \citep{Pedregosa2012}.

The ML method generates ``prediction intervals'' between observed and derived quantities (rather than ``confidence intervals''). A prediction interval is an estimate of the interval encompassing future observations, with a certain probability. And, unlike confidence intervals, which describe uncertainties on the different moments of a population, a prediction interval is unique to each prediction. In many regresson analyses, such as linear fitting, the prediction intervals are based on underlying assumptions of normally distributed residuals. However, RF estimators do not have any such assumptions and require special treatment.

The prediction intervals here are based on the general method of quantile regression forests \citep{Meinshausen2006}. The general idea is that all response variables are recorded, not just the mean. Then the prediction can be returned as the full conditional probability distribution of all responses, which allows us to generate the prediction intervals. We report the 68\% prediction interval as the square root of the second moment of the full conditional probability distribution.
